---
title: "Data Manipulation"
date: 2019-09-19
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import datasets

```{r}
litters_data = read.csv("./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```

## Selecting 

Now we are using the select function to select a subset of columns that we're interested in. First, we have to specify the dataframe followed by the column names. R generates the output in the order of columns we select. 

Selecting does not do anything to the dataframe unless you create a new variable that uses the select function. 

gd0_weight:pups_born_alive keeps all the columns starting with gd0_weight up to pups_born_alive, and gets rid of everything else 

```{r}
select(litters_data, group, litter_number)

select(litters_data, group, litter_number, gd0_weight)

select(litters_data,litter_number, gd0_weight, starts_with("pups"))

select(litters_data, litter_number, group, gd0_weight)

select(litters_data, litter_number, group, everything())

select(litters_data, -group)

select(litters_data, litter_number, gd0_weight:pups_born_alive)

select(litters_data, GROUP = group, litter_number)

rename(litters_data, GROUP = group)
```

## Filter

The == asks whether a variable is a certain value (it's a logical question).

```{r}
filter(litters_data, group == "Mod8")

filter(litters_data, gd_of_birth == 20)

filter(litters_data, gd_of_birth < 20)

filter(litters_data, gd_of_birth >= 20)

filter(litters_data, pups_born_alive < 6)

#filter(litters_data, group == "Con7" | group == "Mod8")
filter(litters_data, group %in% c("Con7", "Mod8"))

filter(litters_data, pups_born_alive >= 4, pups_born_alive <= 6)

filter(litters_data, gd0_weight + gd18_weight < 70)

# what to do about missing values
filter(litters_data, is.na(gd0_weight))

# this will only give us complete rows
drop_na(litters_data)
```

## Mutate

Used to edit variables in the dataset

```{r}
# create wt_gain variable to see how much weight is gained over those 18 days
mutate(
  litters_data, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group),
  group = str_to_upper(group)
)

# str_to_lower makes the variable lowercase
```

## Arrange

```{r}
# puts dataset in order of pup_born_alive 
arrange(litters_data, pups_born_alive)

# arrange in descending order
arrange(litters_data, desc(pups_born_alive))

# arranges data first by pups_born_alive and then within gd0_weight 
arrange(litters_data, pups_born_alive, gd0_weight)
```

## Pipe 

Create a collection of commands

```{r}
litters_data = 
  read.csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight, 
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)
```

```{r}
litters_data = 
  read.csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names(dat = .) %>% 
  select(.data = ., -pups_survive)
```

The . refers back to what was created in the last operaiton